The truncated KdV framework for modeling anomalous waves induced by abrupt depth changes

### 2019-06-25 WavesData-2019/LatexNotes/notes.tex
NM

% Experimental parameters
\newcommand{\lamm}{\lambda^{-}}
\newcommand{\lamp}{\lambda^{+}}
\newcommand{\hm}{h^{-}}
\newcommand{\hp}{h^{+}}
\newcommand{\etastd}{\eta_{\text std}}
% Theory for truncated KdV and stat mech
\newcommand{\uhat}{\hat{u}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\Real}{\text{Re}}
\newcommand{\Fspace}{\mathscr{F}_{\Lambda}}
\newcommand{\Proj}{P_{\Lambda}}
\newcommand{\sumk}{\sum_{k=1}^{\Lambda}}
\newcommand{\sumn}{\sum_{n=1}^{N}}
\newcommand{\uDir}{u_{\text{Dir}}}
\newcommand{\Gibbs}{\mathcal{G}}
\newcommand{\ampp}{\mathcal{A}}
\newcommand{\lamfac}{N}

% New
\newcommand{\Hp}{H^{+}}
\newcommand{\Hm}{H^{-}}
\newcommand{\Gp}{\Gibbs^{+}}
\newcommand{\Gm}{\Gibbs^{-}}
\newcommand{\meanp}[1]{\mean{#1}_{+}}
\newcommand{\meanm}[1]{\mean{#1}_{-}}
\newcommand{\skw}{\text{skew}}
\newcommand{\var}{\text{var}}

% New
\newcommand{\depth}{d}
\newcommand{\dpm}{\depth^{-}}
\newcommand{\dpp}{\depth^{+}}

@@

\subsection{KdV with depth change}
Consider waves propagating unidirectionally in shallow water. Consider the surface displacement $\eta(x,t)$ and the reference frame moving with the local wave speed $\xi = x - ct$, where $c = \sqrt{g \depth}$ is the wave speed, $g$ gravity, and $\depth$ the local depth.
To first-correction in small amplitude, surface displacements are governed by the Korteweg–de Vries equation, which in dimensional form is given by
\begin{equation}
2 \eta_t + \frac{3 c}{\depth} \eta \eta_{\xi} + \frac{c \depth^2}{3} \eta_{\xi \xi \xi} = 0
\end{equation}

We will primarily consider the case in which waves originate from a region of constant depth, encounter an abrupt depth change, and then continue in another region of constant depth. Thus, depth will be piecewise constant
\begin{align}
\depth = 
\begin{cases}
\it \dpm \quad \mbox{if } x<0 \\
\it \dpp \quad \mbox{if } x>0
\end{cases}
\end{align}
Most often, we will consider waves moving into shallower depth, so that $\dpp < \dpm$.

\subsection{Nondimensionalization}

We introduce the following characteristic scales
\begin{align}
\mbox{\em characteristic wave amplitude} \qquad
a = \etastd = \mean{\eta^2}^{1/2} \\
\mbox{\em characteristic wave length} \qquad
 = \etastd = \mean{\eta^2}^{1/2}
\end{align}

>>>>

% New
\newcommand{\etastd}{\eta_{\text std}}
\newcommand{\ampp}{\mathcal{A}}
\newcommand{\Gibbs}{\mathcal{G}}
\newcommand{\lamfac}{N}
% New
\newcommand{\Hp}{H^{+}}
\newcommand{\Hm}{H^{-}}
\newcommand{\Gp}{\Gibbs^{+}}
\newcommand{\Gm}{\Gibbs^{-}}
\newcommand{\meanp}[1]{\mean{#1}_{+}}
\newcommand{\meanm}[1]{\mean{#1}_{-}}
\newcommand{\skw}{\text{skew}}
\newcommand{\var}{\text{var}}

% New
\newcommand{\depth}{d}
\newcommand{\dup}{\depth^{-}}
\newcommand{\ddn}{\depth^{+}}
\newcommand{\lamup}{\lambda^{-}}
\newcommand{\lamdn}{\lambda^{+}}
\newcommand{\freqp}{f_p}

@@

\subsection{The Korteweg–de Vries equation with variable depth}
Consider waves propagating unidirectionally in shallow water. Consider the surface displacement $\eta(x,t)$ and the reference frame moving with the local wave speed $\xi = x - ct$, where $c = \sqrt{g \depth}$ is the wave speed, $g$ gravity, and $\depth$ the local depth.
To first-correction in small amplitude, surface displacements are governed by the Korteweg–de Vries equation (KdV), which in dimensional form is given by
\begin{equation}
2 \eta_t + \frac{3 c}{\depth} \eta \eta_{\xi} + \frac{c \depth^2}{3} \eta_{\xi \xi \xi} = 0
\end{equation}

We will primarily consider the case in which waves originate from a region of constant depth, encounter an abrupt depth change, and then continue in another region of constant depth. Thus, depth will be piecewise constant
\begin{align}
\depth = 
\begin{cases}
\it \dup \quad \mbox{if } x<0 \\
\it \ddn \quad \mbox{if } x>0
\end{cases}
\end{align}
Most often, we will consider waves moving into shallower depth, so that $\ddn < \dup$. The incoming waves are randomized and generated with a peak forcing frequency of $\freqp$.

\subsection{Nondimensionalization of KdV}

We introduce the following characteristic scales
\begin{align}
&a = \etastd = \mean{\eta^2}^{1/2} 
&&\mbox{\em upstream characteristic wave amplitude} \\
&\lamup = \sqrt{g \dup} / \freqp
&&\mbox{\em upstream characteristic wave length}
\end{align}
